Improving the Performance of Low Bandwidth Distributed Data Parallelism

### 2022-12-29 paper/paper.tex
updates to paper

\documentclass[11pt]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{amsmath, amsfonts}
\title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
\author{Christopher Rae \\ \begin{small} email: 110641877@ea.edin.sch.uk \end{small}}

@@

\begin{document}
\maketitle

@@

\section{Methodology}
\subsection*{Distributed Data Parallelism}
The goal of distributed data parallelism is to split the data across the GPUs and computers in the cluster, so that the training process can be completed faster by leveraging the additional computational resources.

@@

\subsection*{SGD}

The training data is denoted as $\{x_i\}^n_{i=1} \subset \mathbb{R}^d$, and the model parameter vector is donated as\\ $w \in \mathbb{R}^d$. The goal of training is to solve the following optimization problem 
$$
\min_w f
(w) = \frac{1}{n} \sum^n_{i=1}L(w,x_i)
$$
We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above. In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} f(w)$. SGD written as:

@@

\subsection*{Gradient Quantization}
Gradient quantization is a technique used in distributed data parallelism to reduce the amount of data that needs to be transmitted over channels during training. It works by reducing the precision of the gradients that are transmitted between worker nodes.

@@

\subsubsection*{Formula}

\subsection*{Gradient Sparsification}
Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting. The goal is to reduce the size of the gradients without significantly affecting their quality, so that they can be transmitted more efficiently over the network.

@@

\subsubsection*{Formula}
We denote sparsified gradients as $S(g, Z_t)$. We have to remember that $g \in \mathbb{R}^d$ so we let $g_i$ be the $i$-th component of the vector ($g=[g_1,\ldots,g_d]$). $Z$ is a vector with the same shape a $g$, when the $w$ is initialized at the start of the training process $Z_i = 0$. The function $S$ has 2 outputs the first being the new representation of the gradient of the weights at $t$, $v$ and the second being the accumulated gradients $Z_{t+1}$

@@

\subsection*{Larger Batch Sizes}
Increasing the batch size in a distributed training setup can also improve training efficiency by reducing the frequency of the gradient averaging step. This is because each device will process a larger amount of data before the gradients are averaged and applied to the model weights. This can reduce the overhead associated with performing the gradient averaging step, and can allow the model to make more efficient use of the available computation resources.

However, it is important to carefully tune the batch size, as increasing it too much can negatively impact the convergence and accuracy of the model. This is because a larger batch size can lead to a more noisy gradient estimate, which can make it more difficult for the model to learn effectively.

>>>>

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{amsmath, amsfonts,tabularx}
\title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
\author{Christopher Rae \\ \begin{small} email: 110641877@ea.edin.sch.uk \end{small}}

@@

\begin{document}
\maketitle
\tableofcontents
\newpage

@@

\section{Methodology}
\subsection{Distributed Data Parallelism}
The goal of distributed data parallelism is to split the data across the GPUs and computers in the cluster, so that the training process can be completed faster by leveraging the additional computational resources.

@@

\subsection{SGD}

The training data is denoted as $\{x_i\}^n_{i=1} \subset \mathbb{R}^d$, and the model parameter vector is donated as\\ $w \in \mathbb{R}^d$. The goal of training is to solve the following optimization problem 
$$
\min_w f
(w) = \frac{1}{n} \sum^n_{i=1}L(w,x_i)
$$
Where $L$ is the loss function. We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above. In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} f(w)$. SGD written as:

@@

\subsection{Gradient Quantization}
Gradient quantization is a technique used in distributed data parallelism to reduce the amount of data that needs to be transmitted over channels during training. It works by reducing the precision of the gradients that are transmitted between worker nodes.

@@

\subsubsection{Formula}

\subsection{Gradient Sparsification}
Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting. The goal is to reduce the size of the gradients without significantly affecting their quality, so that they can be transmitted more efficiently over the network.

@@

\subsubsection{Formula}
We denote sparsified gradients as $S(g, Z_t)$. We have to remember that $g \in \mathbb{R}^d$ so we let $g_i$ be the $i$-th component of the vector ($g=[g_1,\ldots,g_d]$). $Z$ is a vector with the same shape a $g$, when the $w$ is initialized at the start of the training process $Z_i = 0$. The function $S$ has 2 outputs the first being the new representation of the gradient of the weights at $t$, $v$ and the second being the accumulated gradients $Z_{t+1}$

@@

\subsection{Larger Batch Sizes}
In machine learning, the batch size refers to the number of samples that will be processed before the model's weights are updated. The batch size is a hyperparameter that can be chosen when training a model. When using a batch size the SGD algorithm becomes mini-batch SGD it looks something like this:

$$
g_t(w_t) = \frac{1}{B_k} \sum_{i=i}^{B_k} {\nabla} f(w_t)
$$
$$
w_{t+1} = w_t - \eta g_t(w_t)
$$

Increasing the batch size in a distributed training setup can also improve training efficiency by reducing the frequency of the gradient averaging step. This is because each device will process a larger amount of data before the gradients are averaged and applied to the model weights. This can reduce the overhead associated with performing the gradient averaging step, and can allow the model to make more efficient use of the available computation resources.

One negative effect of large batch sizes is that they can lead to a decrease in generalization performance. This occurs because large batch sizes can result in an increase in the variance of the stochastic gradient descent (SGD) updates, which can cause the model to overfit to the training data. This overfitting can manifest as a decrease in performance on held-out test data or a decrease in the ability of the model to generalize to unseen data.

Another negative effect of large batch sizes is that they can lead to a decrease in the convergence rate of the model. This occurs because large batch sizes can result in a reduction in the magnitude of the SGD updates, which can slow down the convergence of the model. In some cases, this can result in the model failing to converge at all.

Furthermore, using large batch sizes can lead to an increase in the amount of memory required to store and process the data. This can be a problem for systems with limited memory or for models with a large number of parameters, as the large batch sizes may not fit in memory.

There are a few things we can do to recude the negative effects of a large batch size.

\subsection{Tests}
\subsubsection{Models}
Within our test we use a range of diferrent models. The most important factor when choosing a model is to make sure we ge a good range of sizes between out models. These are the models we have tested:

\begin{center}
\begin{tabular}{| p{3cm} | p{3cm} | p{3cm} | p{3cm} |} 
\hline
Model & Parameters & Dataset & Task \\ \hline
RESNET-50 & 25,583,592 & ImageNet-1k & MoCo v2 \\ \hline
VGG16 &  138,357,544 & CIFAR-100 & None \\ \hline
Inception v4 & 42,711,400 & ImageNet-1k & MoCo v2\\ \hline
Transformer & 0 & Multiple & GLUE \\ \hline
RoBERTa-Large & 0 & Multiple & GLUE \\ \hline
\end{tabular}
\end{center}
\subsubsection{testing methods}
We start
